---
title: "FRED"
author: "Tanel"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
if (!require("pacman")) install.packages("pacman"); library(pacman)
pacman::p_load(tidyverse,
stringr,
Matrix,
iNEXT,
ggiNEXT,
car,
lmerTest,
ggpubr,
ggplot2,
multcomp,
MuMIn,
reshape2,
car,
nlme,
DHARMa,
gridExtra,
factoextra,
spaMM,
vegan,
magrittr,
rcompanion,
corrr,
GGally,
plotly,
rnaturalearthhires,
cowplot,
sjPlot,
partR2,
sf)
library(hillR)
library(rnaturalearthhires)
library(rnaturalearth)
```


# set wd
```{r}
setwd("C:/Users/pcadmin/Nextcloud/FredAMF")
```

#import data

```{r}
ecobank <- read.csv("C:/Users/pcadmin/Nextcloud/FredAMF/ecobank_full.csv", sep=";")
vt <- read.csv("C:/Users/pcadmin/Nextcloud/FredAMF/ecobank_cultured.csv", sep=";")
fred<- read.csv("C:/Users/pcadmin/Nextcloud/FredAMF/FRED3_fineRoots.csv", sep=",")
```


# order ecobank by sample

```{r}
ecobank <- ecobank[order(ecobank$sample),]
```

# order vt byt sample

```{r}
vt <- vt[order(vt$sample),]
```

# merge ecobank and vt by sample

```{r}
roots <- merge(ecobank, vt, by = "sample",all.x=FALSE)
```

# filter ecobank data where isRoots=TRUE

```{r}
roots <- subset(roots, isRoot == TRUE)
```

# extract unique plant_species

```{r}
plant_species <- unique(roots$plant_species)
plant_species<-as.data.frame(plant_species)
```

# write csv with plant_species

```{r}
write.csv(plant_species, "C:/Users/pcadmin/Nextcloud/FredAMF/plant_species_Tartu.csv")
```


# map points in roots by plot.latitude and plot.longitude in rnaturalearth

# Variable maps
```{r}
basemap <- st_as_sf(ne_countries(scale = 50, returnclass= "sf"))
basemap %<>% st_transform("+proj=longlat +ellps=WGS84")
```


# make first row as header in fred

```{r}
colnames(fred) <- fred[1,]
fred <- fred[-1,]
#rename Latitude_Main to lat and Longitude_Main to lon
fred <- fred %>% rename(lat = Latitude_Main, lon = Longitude_Main)
```



# plot map with points from fred
```{r}
point_data <- fred %>%
  dplyr::select(lon, lat) %>%
  st_as_sf(coords = c("lon", "lat"), crs = "+proj=longlat +ellps=WGS84")

m_altitude <- ggplot(basemap) +
  geom_sf() +
  geom_sf(data = point_data, aes(), alpha = 1, size = 0.7) +
  theme_void()

m_altitude
# save as 30x30 cm png
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/map_fred.jpg", width = 30, height = 15, units = "cm")
```

# plot both roots and fred points on one map
```{r}
point_data <- roots %>% dplyr::select(lon, lat) %>%
  st_as_sf(coords = c("lon", "lat"),
           crs = "+proj=longlat +ellps=WGS84")
point_data_fred <- fred %>%
  dplyr::select(lon, lat) %>%
  st_as_sf(coords = c("lon", "lat"), crs = "+proj=longlat +ellps=WGS84")
m_fredamf <- ggplot(basemap) +
  geom_sf() +
  geom_sf(data = point_data, aes(),color="blue", alpha = 1, size = 0.7) +
  scale_color_viridis_c(name = "richness") +
  geom_sf(data = point_data_fred, aes(),color="red", alpha = 1, size = 0.7) +
  theme_void()
m_fredamf
# save as 30x30 cm png
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/map_both.jpg", width = 30, height = 15, units = "cm")
```

# combine fred columns 18 and 19 into one

```{r}
fred$species <- paste(fred$`Plant taxonomy_Genus_Data Source`, fred$`Plant taxonomy_Species_Data source`, sep = " ")
```

# count unique species in fred
```{r}
unique_fred<-unique(fred$species)
```

# see how many unique overlap with roots$plant_species
```{r}
overlap <- unique_fred[unique_fred %in% roots$plant_species]
length(overlap)
```
#COUNT UNIQUE SPECIES IN ROOTS
```{r}
unique_roots<-unique(roots$plant_species)
```

#overlap of unique_fred and unique_roots

```{r}
overlap <- unique_fred[unique_fred %in% unique_roots]
length(overlap)
```
# make a dataset with only the species that are in both roots and fred
```{r}
fred_roots <- fred[fred$species %in% overlap,]
roots_vt<- roots[roots$plant_species %in% overlap,]
roots_vt$prop.cult<-as.numeric(roots_vt$prop.cult)
roots_vt$prop.ancestral<-as.numeric(roots_vt$prop.ancestral)
roots_vt$prop.edaphophilic<-as.numeric(roots_vt$prop.edaphophilic)
roots_vt$prop.rhizophilic<-as.numeric(roots_vt$prop.rhizophilic)
roots_vt$spec.ances<-as.numeric(roots_vt$spec.ances)
roots_vt$spec.edaph<-as.numeric(roots_vt$spec.edaph)
roots_vt$spec.rhizo<-as.numeric(roots_vt$spec.rhizo)
```

# create a dataset with plant_species by mean richness_raw, prop.cultured, prop.ancestral, prop.edaphophilic, prop.rhizophilic, spec.ances, spec.edaph, spec.rhizo
```{r}
richness_vt <- roots_vt %>%
  group_by(plant_species) %>%
  summarise(mean_richness = mean(richness, na.rm = TRUE),
            mean_prop_cult = mean(prop.cult, na.rm = TRUE),
            mean_prop_ancestral = mean(prop.ancestral, na.rm = TRUE),
            mean_prop_edaphophilic = mean(prop.edaphophilic, na.rm = TRUE),
            mean_prop_rhizophilic = mean(prop.rhizophilic, na.rm = TRUE),
            mean_sp_ancestral = mean(spec.ances,na.rm=TRUE),
            mean_sp_rhizophilic = mean(spec.rhizo,na.rm=TRUE),
            mean_sp_edaphophilic = mean(spec.edaph,na.rm=TRUE)) %>%
  arrange(desc(mean_richness))
```

# create a dataset with species by mean 'Specific root length (SRL)' from fred_roots
```{r}
fred_roots$`Specific root length (SRL)`<- as.numeric(fred_roots$`Specific root length (SRL)`)
srl_fred <- fred_roots %>%
  group_by(species) %>%
  summarise(mean_srl = mean(`Specific root length (SRL)`, na.rm = TRUE)) %>%
  arrange(desc(mean_srl))
```

# do the same for 'Root tissue density (RTD)' from fred_roots
```{r}
fred_roots$`Root tissue density (RTD)`<- as.numeric(fred_roots$`Root tissue density (RTD)`)
rtd_fred <- fred_roots %>%
  group_by(species) %>%
  summarise(mean_rtd = mean(`Root tissue density (RTD)`, na.rm = TRUE)) %>%
  arrange(desc(mean_rtd))
```

# do the same for 'Root diameter (RD)' from fred_roots
```{r}
fred_roots$`Root diameter`<- as.numeric(fred_roots$`Root diameter`)
rd_fred <- fred_roots %>%
  group_by(species) %>%
  summarise(mean_rd = mean(`Root diameter`, na.rm = TRUE)) %>%
  arrange(desc(mean_rd))
```

# do the same for 'Root N content'

```{r}
fred_roots$`Root N content`<- as.numeric(fred_roots$`Root N content`)
rnc_fred <- fred_roots %>%
  group_by(species) %>%
  summarise(mean_rnc = mean(`Root N content`, na.rm = TRUE)) %>%
  arrange(desc(mean_rnc))
```

# merge by species
```{r}
richness_srl_rtd_rd <- merge(richness_vt, srl_fred, by.x = "plant_species", by.y = "species")
richness_srl_rtd_rd <- merge(richness_srl_rtd_rd, rtd_fred, by.x = "plant_species", by.y = "species")
richness_srl_rtd_rd <- merge(richness_srl_rtd_rd, rd_fred, by.x = "plant_species", by.y = "species")
richness_srl_rtd_rd <- merge(richness_srl_rtd_rd, rnc_fred, by.x = "plant_species", by.y = "species")
richness_srl_rtd_rd <- richness_srl_rtd_rd %>%
  arrange(desc(mean_richness))
```

# plot richness vs srl

```{r}
p1<-ggplot(richness_srl_rtd_rd, aes(x = mean_richness, y = mean_srl)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(8, 37) +
  labs(x = "Mean VT richness", y = "Mean SRL(34)") +
  theme_minimal()
```
# plot richness vs rtd
```{r}
p2<-ggplot(richness_srl_rtd_rd, aes(x = mean_richness, y = mean_rtd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(8, 37) +
  labs(x = "Mean VT richness", y = "Mean RTD (30)") +
  theme_minimal()
```
# plot richness vs rd
```{r}
p3<-ggplot(richness_srl_rtd_rd, aes(x = mean_richness, y = mean_rd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(8, 37) +
  labs(x = "Mean VT richness", y = "Mean RD (32)") +
  theme_minimal()
```

# plot richness vs rnc
```{r}
p4<-ggplot(richness_srl_rtd_rd, aes(x = mean_richness, y = mean_rnc)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(8, 37) +
  labs(x = "Mean VT richness", y = "Mean RNC (24)") +
  theme_minimal()
```

# plot all three together
```{r}
library(gridExtra)
p5<-grid.arrange(p1, p2, p3,p4, ncol=2)
```
# save p4 as 30x30 cm png
```{r}
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/richness_srl_rtd_rd.jpg", plot = p5, width = 30, height = 30, units = "cm")
```

# plot prop.cultured vs srl
```{r}
p1<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_prop_cult), y = mean_srl)) +
  geom_point() +
  xlim(-2.5, 0.5) +
  geom_smooth(method = "lm") +
  labs(x = "logit Prop. cultured vt", y = "Mean SRL") +
  theme_minimal()
```
# plot prop.cultured vs rtd
```{r}
p2<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_prop_cult), y = mean_rtd)) +
  geom_point() +
  xlim(-2.5, 0.5) +
  geom_smooth(method = "lm") +
  labs(x = "logit Prop. cultured vt", y = "Mean RTD") +
  theme_minimal()
```

# plot prop.cultured vs rd
```{r}
p3<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_prop_cult), y = mean_rd)) +
  geom_point() +
  xlim(-2.5, 1) +
  geom_smooth(method = "lm") +
  labs(x = "logit Prop. cultured vt", y = "Mean RD") +
  theme_minimal()
```

# plot prop.cultured vs rnc
```{r}
p4<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_prop_cult), y = mean_rnc)) +
  geom_point() +
  xlim(-2.5, 0) +
  geom_smooth(method = "lm") +
  labs(x = "logit Prop. cultured vt", y = "Mean RNC") +
  theme_minimal()
```

# plot all three together
```{r}
p5<-grid.arrange(p1, p2, p3,p4, ncol=2)
```
# save p5 as 30x30 cm png
```{r}
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/prop_cult_srl_rtd_rd.jpg", plot = p5, width = 30, height = 30, units = "cm")
```

# plot srl vs rtd
```{r}
ggplot(richness_srl_rtd_rd, aes(x = mean_srl, y = mean_rtd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Mean SRL", y = "Mean RTD") +
  theme_minimal()
```
# plot srl vs rd
```{r}
ggplot(richness_srl_rtd_rd, aes(x = mean_srl, y = mean_rd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Mean SRL", y = "Mean RD") +
  theme_minimal()
```
# plot prop.ancestral vs srl
```{r}
p1<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_prop_ancestral), y = mean_srl)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "logit Prop. ancestral vt", y = "Mean SRL") +
  theme_minimal()
```
# plot prop.ancestral vs rtd
```{r}
p2<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_prop_ancestral), y = mean_rtd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "logit Prop. ancestral vt", y = "Mean RTD") +
  theme_minimal()
```
# plot prop.ancestral vs rd
```{r}
p3<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_prop_ancestral), y = mean_rd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "logit Prop. ancestral vt", y = "Mean RD") +
  theme_minimal()
```
# plot prop.ancestral vs rnc
```{r}
p4<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_prop_ancestral), y = mean_rnc)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "logit Prop. ancestral vt", y = "Mean RNC") +
  theme_minimal()
```
# plot all three together
```{r}
p5<-grid.arrange(p1, p2, p3,p4, ncol=2)
```
# save p5 as 30x30 cm png
```{r}
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/prop_ancestral_srl_rtd_rd.jpg", plot = p5, width = 30, height = 30, units = "cm")
```

# plot prop.edaphophilic vs srl
```{r}
p1<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_prop_edaphophilic), y = mean_srl)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "logit Prop. edaphophilic vt", y = "Mean SRL") +
  theme_minimal()
```
# plot prop.edaphophilic vs rtd
```{r}
p2<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_prop_edaphophilic), y = mean_rtd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "logit Prop. edaphophilic vt", y = "Mean RTD") +
  theme_minimal()
```
# plot prop.edaphophilic vs rd
```{r}
p3<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_prop_edaphophilic), y = mean_rd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "logit Prop. edaphophilic vt", y = "Mean RD") +
  theme_minimal()
```
# plot prop.edaphophilic vs rnc
```{r}
p4<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_prop_edaphophilic), y = mean_rnc)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(-3.75, -3) +
  labs(x = "logit Prop. edaphophilic vt", y = "Mean RNC") +
  theme_minimal()
```
# plot all three together
```{r}
p5<-grid.arrange(p1, p2, p3,p4, ncol=2)
```

# save p5 as 30x30 cm png
```{r}
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/prop_edaphophilic_srl_rtd_rd.jpg", plot = p5, width = 30, height = 30, units = "cm")
```

# plot prop.rhizophilic vs srl
```{r}
p1<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_prop_rhizophilic), y = mean_srl)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(1.5, 3.8) +
  labs(x = "logit Prop. rhizophilic vt", y = "Mean SRL") +
  theme_minimal()
```
# plot prop.rhizophilic vs rtd
```{r}
p2<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_prop_rhizophilic), y = mean_rtd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(1.5, 3.8) +
  labs(x = "logit Prop. rhizophilic vt", y = "Mean RTD") +
  theme_minimal()
```
# plot prop.rhizophilic vs rd
```{r}
p3<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_prop_rhizophilic), y = mean_rd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(1.5, 3.8) +
  labs(x = "logit Prop. rhizophilic vt", y = "Mean RD") +
  theme_minimal()
```
# plot prop.rhizophilic vs rnc
```{r}
p4<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_prop_rhizophilic), y = mean_rnc)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(1.5, 3.8) +
  labs(x = "logit Prop. rhizophilic vt", y = "Mean RNC") +
  theme_minimal()
```
# plot all three together
```{r}
p5<-grid.arrange(p1, p2, p3,p4, ncol=2)
```
# save p5 as 30x30 cm png
```{r}
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/prop_rhizophilic_srl_rtd_rd.jpg", plot = p5, width = 30, height = 30, units = "cm")
```

# repeat previous with mean_sp_ancestral, mean_sp_rhizophilic, mean_sp_edaphophilic
# plot mean_sp_ancestral vs srl
```{r}
p1<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_sp_ancestral), y = mean_srl)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "logit prop.ances pres/abs", y = "Mean SRL") +
  theme_minimal()
```
# plot mean_sp_ancestral vs rtd
```{r}
p2<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_sp_ancestral), y = mean_rtd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "logit prop.ances pres/abs", y = "Mean RTD") +
  theme_minimal()
```
# plot mean_sp_ancestral vs rd
```{r}
p3<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_sp_ancestral), y = mean_rd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "logit prop.ances pres/abs", y = "Mean RD") +
  theme_minimal()
```
# plot mean_sp_ancestral vs rnc
```{r}
p4<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_sp_ancestral), y = mean_rnc)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "logit prop.rhizo pres/abs", y = "Mean RNC") +
  theme_minimal()
```
# plot all three together
```{r}
p5<-grid.arrange(p1, p2, p3,p4, ncol=2)
```
# save p5 as 30x30 cm png
```{r}
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/mean_sp_ancestral_srl_rtd_rd.jpg", plot = p5, width = 30, height = 30, units = "cm")
```

# plot mean_sp_rhizophilic vs srl
```{r}
p1<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_sp_rhizophilic), y = mean_srl)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "logit prop.rhizo pres/abs", y = "Mean SRL") +
  theme_minimal()
```
# plot mean_sp_rhizophilic vs rtd
```{r}
p2<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_sp_rhizophilic), y = mean_rtd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "logit prop.rhizo pres/abs", y = "Mean RTD") +
  theme_minimal()
```
# plot mean_sp_rhizophilic vs rd
```{r}
p3<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_sp_rhizophilic), y = mean_rd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "logit prop.rhizo pres/abs", y = "Mean RD") +
  theme_minimal()
```
# plot mean_sp_rhizophilic vs rnc
```{r}
p4<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_sp_rhizophilic), y = mean_rnc)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "logit prop.rhizo pres/abs", y = "Mean RNC") +
  theme_minimal()
```
# plot all three together
```{r}
p5<-grid.arrange(p1, p2, p3,p4, ncol=2)
```
# save p5 as 30x30 cm png
```{r}
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/mean_sp_rhizophilic_srl_rtd_rd.jpg", plot = p5, width = 30, height = 30, units = "cm")
```
# plot mean_sp_edaphophilic vs srl
```{r}
p1<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_sp_edaphophilic), y = mean_srl)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "logit prop.edaph pres/abs", y = "Mean SRL") +
  theme_minimal()
```
# plot mean_sp_edaphophilic vs rtd
```{r}
p2<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_sp_edaphophilic), y = mean_rtd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "logit prop.edaph pres/abs", y = "Mean RTD") +
  theme_minimal()
```
# plot mean_sp_edaphophilic vs rd
```{r}
p3<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_sp_edaphophilic), y = mean_rd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "logit prop.edaph pres/abs", y = "Mean RD") +
  theme_minimal()
```
# plot mean_sp_edaphophilic vs rnc
```{r}
p4<-ggplot(richness_srl_rtd_rd, aes(x = logit(mean_sp_edaphophilic), y = mean_rnc)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "logit prop.edaph pres/abs", y = "Mean RNC") +
  theme_minimal()
```
# plot all three together
```{r}
p5<-grid.arrange(p1, p2, p3,p4, ncol=2)
```
# save p5 as 30x30 cm png
```{r}
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/mean_sp_edaphophilic_srl_rtd_rd.jpg", plot = p5, width = 30, height = 30, units = "cm")
```

#import MaarjAM data

```{r}
maarjam <- read.csv("C:/Users/pcadmin/Nextcloud/FredAMF/maarjam_export_host_plants.csv", sep=";")
```

#remove rows from maarjam where species is sp.
```{r}
maarjam <- maarjam[!grepl("sp\\.", maarjam$species), ]
```

# lat and lon as.numeric
```{r}
maarjam$lat <- as.numeric(maarjam$lat)
maarjam$lon <- as.numeric(maarjam$lon)
```

# remove maarjam rows with NA in lat or lon
```{r}
maarjam <- maarjam[!is.na(maarjam$lat) & !is.na(maarjam$lon), ]
```

#combine maarjam columns 18 and 19 into one

```{r}
maarjam$species <- paste(maarjam$genus, maarjam$species, sep = " ")
```

# count unique species in maarjam
```{r}
unique_maarjam<-unique(maarjam$species)
```
# see how many unique overlap with roots$plant_species
```{r}
overlap <- unique_maarjam[unique_maarjam %in% roots$plant_species]
length(overlap)
```
# keep only species from overlap in maarjam
```{r}
maarjam <- maarjam[maarjam$species %in% overlap,]
```

# count unique species in maarjam
```{r}
unique_maarjam<-unique(maarjam$species)
```

# see how many unique_fred overlap with unique_maarjam
```{r}
overlap <- unique_fred[unique_fred %in% unique_maarjam]
length(overlap)
```

# plot maarjam and fred on one map, give points in maarjam names by smID
```{r}
point_data <- maarjam %>%
  dplyr::select(lon, lat) %>%
  st_as_sf(coords = c("lon", "lat"), crs = "+proj=longlat +ellps=WGS84")
point_data_fred <- fred %>%
  dplyr::select(lon, lat) %>%
  st_as_sf(coords = c("lon", "lat"), crs = "+proj=longlat +ellps=WGS84")
m_fredamf <- ggplot(basemap) +
  geom_sf() +
  geom_sf(data = point_data, aes(),color="blue", alpha = 1, size = 0.7) +
  scale_color_viridis_c(name = "richness") +
  geom_sf(data = point_data_fred, aes(),color="red", alpha = 1, size = 0.7) +
  theme_void()
m_fredamf
# save as 30x30 cm png
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/map_both_maarjam.jpg", width = 30, height = 15, units = "cm")
```
# create a data table where species are in first column and virtTaxon in other columns

```{r}
# create a data table with maarjam species and their corresponding virtTaxon
# species in first column and virtTaxon in other columns
maarjam_byvt <- maarjam %>%
  group_by(species) %>%
  summarise(virtTaxon = paste(unique(virtTaxon), collapse = ", ")) %>%
  arrange(species)
```

# split maarjam_byvt$virtTaxon into separate columns
```{r}
maarjam_byvt <- maarjam_byvt %>%
  separate(virtTaxon, into = paste0("virtTaxon", 1:148), sep = ", ", fill = "right", extra = "drop")
```
# replace NULL with NA
```{r}
maarjam_byvt[maarjam_byvt == "NULL"] <- NA
```
# remove rows with colums 2:149 that are all NA
```{r}
maarjam_byvt <- maarjam_byvt %>%
  filter(rowSums(is.na(.[2:149])) != ncol(.[2:149]))
```
#write csv file with maarjam_byvt
```{r}
write.csv(maarjam_byvt, "C:/Users/pcadmin/Nextcloud/FredAMF/maarjam_byvt.csv", row.names = FALSE)
```

# create a string with unique from maarjam_byvt columns 2:149
```{r}
unique_virtTaxon <- unique(unlist(maarjam_byvt[2:149]))
```

# write csv file with unique_virtTaxon
```{r}
write.csv(unique_virtTaxon, "C:/Users/pcadmin/Nextcloud/FredAMF/unique_virtTaxon.csv", row.names = FALSE)
```

# import plant_VT_traits

```{r}
maarjam2 <- read.csv("C:/Users/pcadmin/Nextcloud/FredAMF/plant_VT_traits.csv", sep=";")
```

# filter fred to have only species in maarjam2

```{r}
fred2 <- fred[fred$species %in% maarjam2$species,]
```

# count unique species in fred2
```{r}
unique_fred2<-unique(fred2$species)
```

# create a dataset with species by mean 'Specific root length (SRL)' from fred2

```{r}
fred2$`Specific root length (SRL)`<- as.numeric(fred2$`Specific root length (SRL)`)
srl_fred2 <- fred2 %>%
  group_by(species) %>%
  summarise(mean_srl = mean(`Specific root length (SRL)`, na.rm = TRUE)) %>%
  arrange(desc(mean_srl))
```
# create a dataset with species by mean 'Root tissue density (RTD)' from fred2
```{r}
fred2$`Root tissue density (RTD)`<- as.numeric(fred2$`Root tissue density (RTD)`)
rtd_fred2 <- fred2 %>%
  group_by(species) %>%
  summarise(mean_rtd = mean(`Root tissue density (RTD)`, na.rm = TRUE)) %>%
  arrange(desc(mean_rtd))
```
# create a dataset with species by mean 'Root diameter (RD)' from fred2
```{r}
fred2$`Root diameter`<- as.numeric(fred2$`Root diameter`)
rd_fred2 <- fred2 %>%
  group_by(species) %>%
  summarise(mean_rd = mean(`Root diameter`, na.rm = TRUE)) %>%
  arrange(desc(mean_rd))
```
# create a dataset with species by mean 'Root N content' from fred2
```{r}
fred2$`Root N content`<- as.numeric(fred2$`Root N content`)
rnc_fred2 <- fred2 %>%
  group_by(species) %>%
  summarise(mean_rnc = mean(`Root N content`, na.rm = TRUE)) %>%
  arrange(desc(mean_rnc))
```
# merge by species
```{r}
richness_srl_rtd_rd2 <- merge(richness_vt, srl_fred2, by.x = "plant_species", by.y = "species")
richness_srl_rtd_rd2 <- merge(richness_srl_rtd_rd2, rtd_fred2, by.x = "plant_species", by.y = "species")
richness_srl_rtd_rd2 <- merge(richness_srl_rtd_rd2, rd_fred2, by.x = "plant_species", by.y = "species")
richness_srl_rtd_rd2 <- merge(richness_srl_rtd_rd2, rnc_fred2, by.x = "plant_species", by.y = "species")
richness_srl_rtd_rd2 <- richness_srl_rtd_rd2 %>%
  arrange(desc(mean_richness))
```
# plot richness vs srl
```{r}
p1<-ggplot(richness_srl_rtd_rd2, aes(x = mean_richness, y = mean_srl)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(8, 37) +
  labs(x = "Mean VT richness", y = "Mean SRL(34)") +
  theme_minimal()
```
# plot richness vs rtd
```{r}
p2<-ggplot(richness_srl_rtd_rd2, aes(x = mean_richness, y = mean_rtd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(8, 37) +
  labs(x = "Mean VT richness", y = "Mean RTD (30)") +
  theme_minimal()
```
# plot richness vs rd
```{r}
p3<-ggplot(richness_srl_rtd_rd2, aes(x = mean_richness, y = mean_rd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(8, 37) +
  labs(x = "Mean VT richness", y = "Mean RD (32)") +
  theme_minimal()
```
# plot richness vs rnc
```{r}
p4<-ggplot(richness_srl_rtd_rd2, aes(x = mean_richness, y = mean_rnc)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(8, 37) +
  labs(x = "Mean VT richness", y = "Mean RNC (24)") +
  theme_minimal()
```
# plot all three together
```{r}
p5<-grid.arrange(p1, p2, p3,p4, ncol=2)
```
# save p5 as 30x30 cm png
```{r}
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/richness_srl_rtd_rd2.jpg", plot = p5, width = 30, height = 30, units = "cm")
```

# plot prop.cultured vs srl
```{r}
p1<-ggplot(richness_srl_rtd_rd2, aes(x = logit(mean_prop_cult), y = mean_srl)) +
  geom_point() +
  xlim(-2.5, 0.5) +
  geom_smooth(method = "lm") +
  labs(x = "logit Prop. cultured vt", y = "Mean SRL") +
  theme_minimal()
```
# plot prop.cultured vs rtd
```{r}
p2<-ggplot(richness_srl_rtd_rd2, aes(x = logit(mean_prop_cult), y = mean_rtd)) +
  geom_point() +
  xlim(-2.5, 0.5) +
  geom_smooth(method = "lm") +
  labs(x = "logit Prop. cultured vt", y = "Mean RTD") +
  theme_minimal()
```
# plot prop.cultured vs rd
```{r}
p3<-ggplot(richness_srl_rtd_rd2, aes(x = logit(mean_prop_cult), y = mean_rd)) +
  geom_point() +
  xlim(-2.5, 1) +
  geom_smooth(method = "lm") +
  labs(x = "logit Prop. cultured vt", y = "Mean RD") +
  theme_minimal()
```
# plot prop.cultured vs rnc
```{r}
p4<-ggplot(richness_srl_rtd_rd2, aes(x = logit(mean_prop_cult), y = mean_rnc)) +
  geom_point() +
  xlim(-2.5, 0) +
  geom_smooth(method = "lm") +
  labs(x = "logit Prop. cultured vt", y = "Mean RNC") +
  theme_minimal()
```
# plot all three together
```{r}
p5<-grid.arrange(p1, p2, p3,p4, ncol=2)
```


# create new dataset from fred with only columns 'Plant taxonomy_Accepted genus_TPL', 'Specific root length (SRL)', 'Root tissue density (RTD)', 'Root diameter', 'Root N content'

```{r}
fred3 <- fred %>%
  select(`Plant taxonomy_Accepted genus_TPL`, `Specific root length (SRL)`, `Root tissue density (RTD)`, `Root diameter`, `Root N content`, `Plant woodiness_TRY`) %>%
  rename(genus = `Plant taxonomy_Accepted genus_TPL`, srl = `Specific root length (SRL)`, rtd = `Root tissue density (RTD)`, rd = `Root diameter`, rnc = `Root N content`, woodiness = `Plant woodiness_TRY`)
```

# refactor woodiness to woody = 1, non-woody = 0

```{r}
fred3$woodiness <- ifelse(fred3$woodiness == "woody", 1, 0)
```



# create a dataset with genus by mean 'Specific root length (SRL)', 'Root tissue density (RTD)', 'Root diameter', 'Root N content'
```{r}
fred3$srl <- as.numeric(fred3$srl)
fred3$rtd <- as.numeric(fred3$rtd)
fred3$rd <- as.numeric(fred3$rd)
fred3$rnc <- as.numeric(fred3$rnc)
fred_genus <- fred3 %>%
  group_by(genus) %>%
  summarise(mean_srl = mean(srl, na.rm = TRUE),
            mean_rtd = mean(rtd, na.rm = TRUE),
            mean_rd = mean(rd, na.rm = TRUE),
            mean_rnc = mean(rnc, na.rm = TRUE),
            mean_woody = mean(woodiness, na.rm=TRUE)) %>%
  arrange(desc(mean_srl))
```
# boxplots of 'srl', 'rtd', 'rd', 'rnc' by genus. omit NA values
```{r}
p1 <- ggplot(fred3[fred3$srl>0,], aes(x = genus, y = srl)) +
  geom_boxplot() +
  labs(x = "Genus", y = "Specific root length (SRL)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
p2 <- ggplot(fred3[fred3$srl>0,], aes(x = genus, y = rtd)) +
  geom_boxplot() +
  labs(x = "Genus", y = "Root tissue density (RTD)") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
p3 <- ggplot(fred3[fred3$srl>0,], aes(x = genus, y = rd)) +
  geom_boxplot() +
  labs(x = "Genus", y = "Root diameter") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
p4 <- ggplot(fred3[fred3$srl>0,], aes(x = genus, y = rnc)) +
  geom_boxplot() +
  labs(x = "Genus", y = "Root N content") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
# arrange boxplots in a grid
p5 <- grid.arrange(p1, p2, p3, p4, ncol = 2)
# save boxplots as 30x30 cm png
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/genus_srl_rtd_rd_rnc.jpg", plot = p5, width = 45, height = 30, units = "cm")
```

# import maarjam_genus

```{r}
maarjam_genus <- read.csv("C:/Users/pcadmin/Nextcloud/FredAMF/maarjam_genus.csv", sep=";")
```

# count unique genus in maarjam_genus

```{r}
unique_maarjam_genus <- unique(maarjam_genus$genus)
#451
```

# count unique genus in fred_genus

```{r}
unique_fred_genus <- unique(fred_genus$genus)
# 136
```

# see how many unique genus overlap with maarjam_genus
```{r}
overlap_genus <- unique_fred_genus[unique_fred_genus %in% unique_maarjam_genus]
length(overlap_genus)
# 124
```
# keep only genus from overlap in fred_genus
```{r}
fred_genus <- fred_genus[fred_genus$genus %in% overlap_genus,]
```

# keep only overlapping genus in maajram

```{r}
maarjam_genus <- maarjam_genus[maarjam_genus$genus %in% overlap_genus,]
```

# make lists of virtTaxon for each genus in maarjam_genus

```{r}
maarjam_genus_virtTaxon <- maarjam_genus %>%
  group_by(genus) %>%
  summarise(virtTaxon = paste(unique(virtTaxon), collapse = ", ")) %>%
  arrange(genus)
```

# split maarjam_genus_virtTaxon$virtTaxon into separate columns

```{r}
maarjam_genus_virtTaxon <- maarjam_genus_virtTaxon %>%
  separate(virtTaxon, into = paste0("virtTaxon", 1:148), sep = ", ", fill = "right", extra = "drop")
```

# replace NULL with NA
```{r}
maarjam_genus_virtTaxon[maarjam_genus_virtTaxon == "NULL"] <- NA
```

# write csv file with maarjam_genus_virtTaxon
```{r}
write.csv(maarjam_genus_virtTaxon, "C:/Users/pcadmin/Nextcloud/FredAMF/maarjam_genus_virtTaxon.csv", row.names = FALSE)
```

# make a list of all virtTaxon in maarjam_genus_virtTaxon

```{r}
unique_virtTaxon_genus <- unique(unlist(maarjam_genus_virtTaxon[2:149]))
unique_virtTaxon_genus <- as.factor(unique_virtTaxon_genus)
```

# write csv file with unique_virtTaxon_genus
```{r}
write.csv(unique_virtTaxon_genus, "C:/Users/pcadmin/Nextcloud/FredAMF/unique_virtTaxon_genus.csv", row.names = FALSE)
```

# import genus_presabs.csv

```{r}
genus_presabs <- read.csv("C:/Users/pcadmin/Nextcloud/FredAMF/genus_presabs.csv", sep=";")
```

# calculate rowsums and make new column called richness

```{r}
genus_presabs$richness <- rowSums(genus_presabs[2:149], na.rm = TRUE)
```

# merge genus_presabs with fred_genus
```{r}
richness_genus <- merge(fred_genus, genus_presabs, by.x = "genus", by.y = "genus")
richness_genus <- richness_genus %>%
  arrange(desc(richness))
```

# plot richness vs

```{r}
p1 <- ggplot(richness_genus, aes(x = richness, y = mean_srl)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(aes(label = genus), hjust = 0.5, vjust = -0.5, size = 3) +
  xlim(0, 90) +
  labs(x = "VT richness", y = "Mean SRL") +
  theme_minimal()

p2 <- ggplot(richness_genus, aes(x = richness, y = mean_rtd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(aes(label = genus), hjust = 0.5, vjust = -0.5, size = 3) +
  xlim(0, 90) +
  labs(x = "VT richness", y = "Mean RTD") +
  theme_minimal()

p3 <- ggplot(richness_genus, aes(x = richness, y = mean_rd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(aes(label = genus), hjust = 0.5, vjust = -0.5, size = 3) +
  xlim(0, 90) +
  labs(x = "VT richness", y = "Mean RD") +
  theme_minimal()

p4 <- ggplot(richness_genus, aes(x = richness, y = mean_rnc)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(aes(label = genus), hjust = 0.5, vjust = -0.5, size = 3) +
  xlim(0, 90) +
  labs(x = "VT richness", y = "Mean RNC") +
  theme_minimal()

p5 <- grid.arrange(p1, p2, p3, p4, ncol = 2)
```
# save p5 as 30x30 cm png
```{r}
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/richness_genus_srl_rtd_rd_rnc.jpg", plot = p5, width = 35, height = 25, units = "cm")
```

# plot prop_cult vs srl rtd rd rnc
```{r}
p1 <- ggplot(richness_genus, aes(x = logit(prop_cult), y = mean_srl)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(-2, 1) +
  labs(x = "logit Prop. cultured vt", y = "Mean SRL") +
  theme_minimal()
p2 <- ggplot(richness_genus, aes(x = logit(prop_cult), y = mean_rtd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(-2, 1) +
  labs(x = "logit Prop. cultured vt", y = "Mean RTD") +
  theme_minimal()
p3 <- ggplot(richness_genus, aes(x = logit(prop_cult), y = mean_rd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(-2, 1) +
  labs(x = "logit Prop. cultured vt", y = "Mean RD") +
  theme_minimal()
p4 <- ggplot(richness_genus, aes(x = logit(prop_cult), y = mean_rnc)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(-2, 1) +
  labs(x = "logit Prop. cultured vt", y = "Mean RNC") +
  theme_minimal()
p5 <- grid.arrange(p1, p2, p3, p4, ncol = 2)
```
# save p5 as 30x30 cm png
```{r}
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/prop_cult_genus_srl_rtd_rd_rnc.jpg", plot = p5, width = 20, height = 15, units = "cm")
```

# plot prop_rhiz vs srl rtd rd rnc. omit where prop_rhiz is 0

```{r}
p1 <- ggplot(richness_genus[richness_genus$prop_rhiz > 0, ], aes(x = logit(prop_rhiz), y = mean_srl)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "logit Prop. rhizophilic vt", y = "Mean SRL") +
  theme_minimal()
p2 <- ggplot(richness_genus[richness_genus$prop_rhiz > 0, ], aes(x = logit(prop_rhiz), y = mean_rtd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "logit Prop. rhizophilic vt", y = "Mean RTD") +
  theme_minimal()
p3 <- ggplot(richness_genus[richness_genus$prop_rhiz > 0, ], aes(x = logit(prop_rhiz), y = mean_rd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "logit Prop. rhizophilic vt", y = "Mean RD") +
  theme_minimal()
p4 <- ggplot(richness_genus[richness_genus$prop_rhiz > 0, ], aes(x = logit(prop_rhiz), y = mean_rnc)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "logit Prop. rhizophilic vt", y = "Mean RNC") +
  theme_minimal()

p5 <- grid.arrange(p1, p2, p3, p4, ncol = 2)
```
  
# save p5 as 30x30 cm png
```{r}
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/prop_rhiz_genus_srl_rtd_rd_rnc.jpg", plot = p5, width = 20, height = 15, units = "cm")
```

# plot prop_edap vs srl rtd rd rnc. omit where prop_edap is 0
```{r}
p1 <- ggplot(richness_genus[richness_genus$prop_edap > 0, ], aes(x = logit(prop_edap), y = mean_srl)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(-4, -1.5) +
  labs(x = "logit Prop. edaphophilic vt", y = "Mean SRL") +
  theme_minimal()
p2 <- ggplot(richness_genus[richness_genus$prop_edap > 0, ], aes(x = logit(prop_edap), y = mean_rtd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(-4, -1.5) +
  labs(x = "logit Prop. edaphophilic vt", y = "Mean RTD") +
  theme_minimal()
p3 <- ggplot(richness_genus[richness_genus$prop_edap > 0, ], aes(x = logit(prop_edap), y = mean_rd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(-4, -1.5) +
  labs(x = "logit Prop. edaphophilic vt", y = "Mean RD") +
  theme_minimal()
p4 <- ggplot(richness_genus[richness_genus$prop_edap > 0, ], aes(x = logit(prop_edap), y = mean_rnc)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(-4, -1.5) +
  labs(x = "logit Prop. edaphophilic vt", y = "Mean RNC") +
  theme_minimal()
p5 <- grid.arrange(p1, p2, p3, p4, ncol = 2)
```
# save p5 as 30x30 cm png
```{r}
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/prop_edap_genus_srl_rtd_rd_rnc.jpg", plot = p5, width = 20, height = 15, units = "cm")
```

# plot prop_anc vs srl rtd rd rnc. omit where prop_anc is 0

```{r}
p1 <- ggplot(richness_genus[richness_genus$prop_anc > 0, ], aes(x = logit(prop_anc), y = mean_srl)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(-4, -1.5) +
  labs(x = "logit Prop. ancestral vt", y = "Mean SRL") +
  theme_minimal()
p2 <- ggplot(richness_genus[richness_genus$prop_anc > 0, ], aes(x = logit(prop_anc), y = mean_rtd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(-4, -1.5) +
  labs(x = "logit Prop. ancestral vt", y = "Mean RTD") +
  theme_minimal()
p3 <- ggplot(richness_genus[richness_genus$prop_anc > 0, ], aes(x = logit(prop_anc), y = mean_rd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(-4, -1.5) +
  labs(x = "logit Prop. ancestral vt", y = "Mean RD") +
  theme_minimal()
p4 <- ggplot(richness_genus[richness_genus$prop_anc > 0, ], aes(x = logit(prop_anc), y = mean_rnc)) +
  geom_point() +
  geom_smooth(method = "lm") +
  xlim(-4, -1.5) +
  labs(x = "logit Prop. ancestral vt", y = "Mean RNC") +
  theme_minimal()

p5 <- grid.arrange(p1, p2, p3, p4, ncol = 2)
```
# save p5 as 30x30 cm png
```{r}
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/prop_anc_genus_srl_rtd_rd_rnc.jpg", plot = p5, width = 20, height = 15, units = "cm")
```

# PCA of richness_genus with na.omit

```{r}
roots_st <- scale(fred3[, c("srl", "rtd", "rd", "rnc", "woodiness")])
roots_pca <- prcomp(na.omit(roots_st), scale. = FALSE, center = FALSE)
summary(roots_pca)
pca_fig<-fviz_pca_biplot(roots_pca,
label = "var", repel = TRUE)
#save roots_pca as 30x30 cm png
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/roots_pca.jpg", plot = pca_fig, width = 15, height = 15, units = "cm")
```
  
# separate richness_genus into two datasets based on mean_woody = 0 and >0
```{r}
richness_genus_woody <- richness_genus[richness_genus$mean_woody > 0, ]
richness_genus_nonwoody <- richness_genus[richness_genus$mean_woody == 0, ]
```

# PCA of richness_genus with na.omit

```{r}
roots_st <- scale(fred3[, c("srl", "rtd", "rd", "rnc")])
roots_pca <- prcomp(na.omit(roots_st), scale. = FALSE, center = FALSE)
summary(roots_pca)
pca_fig<-fviz_pca_biplot(roots_pca,
label = "var", repel = TRUE)
#save roots_pca as 30x30 cm png
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/roots_pca_4way.jpg", plot = pca_fig, width = 15, height = 15, units = "cm")
```
# pca in richness_genus_woody and richness_genus_nonwoody

```{r}
roots_st_woody <- scale(richness_genus_woody[, c("mean_srl", "mean_rtd", "mean_rd", "mean_rnc")])
roots_pca_woody <- prcomp(na.omit(roots_st_woody), scale. = FALSE, center = FALSE)
summary(roots_pca_woody)
pca_fig_woody <- fviz_pca_biplot(roots_pca_woody,
label = "var", repel = TRUE) +
  labs(title = "PCA of woody genera")
#save roots_pca_woody as 30x30 cm png
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/roots_pca_woody.jpg", plot = pca_fig_woody, width = 15, height = 15, units = "cm")
```

```{r}
roots_st_nonwoody <- scale(richness_genus_nonwoody[, c("mean_srl", "mean_rtd", "mean_rd", "mean_rnc")])
roots_pca_nonwoody <- prcomp(na.omit(roots_st_nonwoody), scale. = FALSE, center = FALSE)
summary(roots_pca_nonwoody)
pca_fig_nonwoody <- fviz_pca_biplot(roots_pca_nonwoody,
label = "var", repel = TRUE) +
  labs(title = "PCA of non-woody genera")
#save roots_pca_nonwoody as 30x30 cm png
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/roots_pca_nonwoody.jpg", plot = pca_fig_nonwoody, width = 15, height = 15, units = "cm")
```

# plot richness vs srl rtd rd rnc for woody and non-woody separately

```{r}
p1_woody <- ggplot(richness_genus_woody, aes(x = richness, y = mean_srl)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(aes(label = genus), hjust = 0.5, vjust = -0.5, size = 3) +
  xlim(0, 90) +
  labs(x = "VT richness", y = "Mean SRL") +
  theme_minimal()
p2_woody <- ggplot(richness_genus_woody, aes(x = richness, y = mean_rtd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(aes(label = genus), hjust = 0.5, vjust = -0.5, size = 3) +
  xlim(0, 90) +
  labs(x = "VT richness", y = "Mean RTD") +
  theme_minimal()
p3_woody <- ggplot(richness_genus_woody, aes(x = richness, y = mean_rd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(aes(label = genus), hjust = 0.5, vjust = -0.5, size = 3) +
  xlim(0, 90) +
  labs(x = "VT richness", y = "Mean RD") +
  theme_minimal()
p4_woody <- ggplot(richness_genus_woody, aes(x = richness, y = mean_rnc)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(aes(label = genus), hjust = 0.5, vjust = -0.5, size = 3) +
  xlim(0, 90) +
  labs(x = "VT richness", y = "Mean RNC") +
  theme_minimal()
p5_woody <- grid.arrange(p1_woody, p2_woody, p3_woody, p4_woody, ncol = 2)
# save p5_woody as 30x30 cm png
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/richness_genus_woody_srl_rtd_rd_rnc.jpg", plot = p5_woody, width = 15, height = 15, units = "cm")
```

```{r}
p1_nonwoody <- ggplot(richness_genus_nonwoody, aes(x = richness, y = mean_srl)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(aes(label = genus), hjust = 0.5, vjust = -0.5, size = 3) +
  xlim(0, 90) +
  labs(x = "VT richness", y = "Mean SRL") +
  theme_minimal()
p2_nonwoody <- ggplot(richness_genus_nonwoody, aes(x = richness, y = mean_rtd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(aes(label = genus), hjust = 0.5, vjust = -0.5, size = 3) +
  xlim(0, 90) +
  labs(x = "VT richness", y = "Mean RTD") +
  theme_minimal()
p3_nonwoody <- ggplot(richness_genus_nonwoody, aes(x = richness, y = mean_rd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(aes(label = genus), hjust = 0.5, vjust = -0.5, size = 3) +
  xlim(0, 90) +
  labs(x = "VT richness", y = "Mean RD") +
  theme_minimal()
p4_nonwoody <- ggplot(richness_genus_nonwoody, aes(x = richness, y = mean_rnc)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_text(aes(label = genus), hjust = 0.5, vjust = -0.5, size = 3) +
  xlim(0, 90) +
  labs(x = "VT richness", y = "Mean RNC") +
  theme_minimal()
p5_nonwoody <- grid.arrange(p1_nonwoody, p2_nonwoody, p3_nonwoody, p4_nonwoody, ncol = 2)
# save p5_nonwoody as 30x30 cm png
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/richness_genus_nonwoody_srl_rtd_rd_rnc.jpg", plot = p5_nonwoody, width = 15, height = 15, units = "cm")
```
# import FR_species.csv (FungalRoot)

```{r}
fr_species <- read.csv("C:/Users/pcadmin/Nextcloud/FredAMF/FR_species.csv", sep=",")
```
# filter to have only rows where mean_col is not NA

```{r}
fr_species <- fr_species %>%
  filter(!is.na(mean_col))
```

# create new dataset from fred with only columns 'Plant taxonomy_Accepted species_TPL', 'Specific root length (SRL)', 'Root tissue density (RTD)', 'Root diameter', 'Root N content'

```{r}
fred4 <- fred %>%
  select(`Plant Taxonomy_Accepted species_TPL`,`Plant taxonomy_Accepted genus_TPL`, `Specific root length (SRL)`, `Root tissue density (RTD)`, `Root diameter`, `Root N content`, `Plant woodiness_TRY`) %>%
  rename(species = `Plant Taxonomy_Accepted species_TPL`, genus = `Plant taxonomy_Accepted genus_TPL`, srl = `Specific root length (SRL)`, rtd = `Root tissue density (RTD)`, rd = `Root diameter`, rnc = `Root N content`, woodiness = `Plant woodiness_TRY`)
```

# combine columns genus and species in fred4

```{r}
fred4$species <- paste(fred4$genus, fred4$species, sep = " ")
fred4 <- fred4 %>%
  select(species, srl, rtd, rd, rnc, woodiness)
```

# merge fred4 with fr_species by species

```{r}
richness_fr_species <- merge(fred4, fr_species, by.x = "species", by.y = "species")
richness_fr_species$srl <- as.numeric(richness_fr_species$srl)
richness_fr_species$rtd <- as.numeric(richness_fr_species$rtd)
richness_fr_species$rd <- as.numeric(richness_fr_species$rd)
richness_fr_species$rnc <- as.numeric(richness_fr_species$rnc)
richness_fr_species$mean_col <- as.numeric(richness_fr_species$mean_col)
richness_fr_species$woodiness <- ifelse(richness_fr_species$woodiness == "woody", 1, 0)
richness_fr_species <- richness_fr_species %>%
  select(species, srl, rtd, rd, rnc, woodiness, mean_col)
```

# mean by species for 'srl', 'rtd', 'rd', 'rnc' and mean_col

```{r}
richness_fr_species <- richness_fr_species %>%
  group_by(species) %>%
  summarise(mean_srl = mean(srl, na.rm = TRUE),
            mean_rtd = mean(rtd, na.rm = TRUE),
            mean_rd = mean(rd, na.rm = TRUE),
            mean_rnc = mean(rnc, na.rm = TRUE),
            mean_col = mean(mean_col, na.rm = TRUE),
            mean_woody = mean(woodiness,na.rm=TRUE)) %>%
  arrange(desc(mean_col))
```

# plot mean_col vs srl rtd rd rnc

```{r}
p1 <- ggplot(richness_fr_species, aes(x = mean_col, y = mean_srl)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Mean col%", y = "Mean SRL") +
  theme_minimal()
p2 <- ggplot(richness_fr_species, aes(x = mean_col, y = mean_rtd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Mean col%", y = "Mean RTD") +
  theme_minimal()
p3 <- ggplot(richness_fr_species, aes(x = mean_col, y = mean_rd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Mean col%", y = "Mean RD") +
  theme_minimal()
p4 <- ggplot(richness_fr_species, aes(x = mean_col, y = mean_rnc)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Mean col%", y = "Mean RNC") +
  theme_minimal()
p5 <- grid.arrange(p1, p2, p3, p4, ncol = 2)
# save p5 as 30x30 cm png
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/richness_fr_species_srl_rtd_rd_rnc.jpg", plot = p5, width = 20, height = 15, units = "cm")
```

# separate woody and non-woody species

```{r}
richness_fr_species_woody <- richness_fr_species[richness_fr_species$mean_woody > 0, ]
richness_fr_species_nonwoody <- richness_fr_species[richness_fr_species$mean_woody == 0, ]
```

# plot mean_col vs srl rtd rd rnc for woody species

```{r}
p1_woody <- ggplot(richness_fr_species_woody, aes(x = mean_col, y = mean_srl)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Mean col%", y = "Mean SRL") +
  theme_minimal()
p2_woody <- ggplot(richness_fr_species_woody, aes(x = mean_col, y = mean_rtd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Mean col%", y = "Mean RTD") +
  theme_minimal()
p3_woody <- ggplot(richness_fr_species_woody, aes(x = mean_col, y = mean_rd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Mean col%", y = "Mean RD") +
  theme_minimal()
p4_woody <- ggplot(richness_fr_species_woody, aes(x = mean_col, y = mean_rnc)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Mean col%", y = "Mean RNC") +
  theme_minimal()
p5_woody <- grid.arrange(p1_woody, p2_woody, p3_woody, p4_woody, ncol = 2)
# save p5_woody as 30x30 cm png
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/richness_fr_species_woody_srl_rtd_rd_rnc.jpg", plot = p5_woody, width = 20, height = 15, units = "cm")
```
```{r}
p1_nonwoody <- ggplot(richness_fr_species_nonwoody, aes(x = mean_col, y = mean_srl)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Mean col%", y = "Mean SRL") +
  theme_minimal()
p2_nonwoody <- ggplot(richness_fr_species_nonwoody, aes(x = mean_col, y = mean_rtd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Mean col%", y = "Mean RTD") +
  theme_minimal()
p3_nonwoody <- ggplot(richness_fr_species_nonwoody, aes(x = mean_col, y = mean_rd)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Mean col%", y = "Mean RD") +
  theme_minimal()
p4_nonwoody <- ggplot(richness_fr_species_nonwoody, aes(x = mean_col, y = mean_rnc)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(x = "Mean col%", y = "Mean RNC") +
  theme_minimal()
p5_nonwoody <- grid.arrange(p1_nonwoody, p2_nonwoody, p3_nonwoody, p4_nonwoody, ncol = 2)
# save p5_nonwoody as 30x30 cm png
ggsave("C:/Users/pcadmin/Nextcloud/FredAMF/richness_fr_species_nonwoody_srl_rtd_rd_rnc.jpg", plot = p5_nonwoody, width = 20, height = 15, units = "cm")
```

# count nr of unique plant_species in ecobank
```{r}
unique_ecobank_species <- unique(ecobank$plant_species)
length(unique_ecobank_species)
```
